---
title: "Forced Displacement Survey - South Sudan"
author: "Data4Decision team - data source : Forced Displacement Survey - South Sudan"
date: "`r format(Sys.time(), '%d %B %Y')`"
always_allow_html: yes
header-includes:
  - \usepackage{booktabs}
  - \usepackage{float}
  - \usepackage{color}
  - \usepackage{longtable}
  - \usepackage{multirow}
output: 
  # prettydoc::html_pretty:
  #    theme: cayman
  officedown::rdocx_document:
    tables:
      style: Table
      layout: autofit
      width: 1
      topcaption: true
      tab.lp: 'tab:'
      caption:
        style: Table Caption
        pre: 'Tab '
        sep: ': '
        tnd: 0
        tns: '-'
        fp_text: !expr officer::fp_text_lite(bold = TRUE)
      conditional:
        first_row: true
        first_column: false
        last_row: false
        last_column: false
        no_hband: false
        no_vband: true
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: true
    toc_depth: 3
    number_sections: true
    theme: lumen
  pdf_document:
    keep_tex: no
  word_document:
    fig_width: 6.5
    fig_height: 3.5
    fig_caption: true
#html_notebook: default
---
  
\newpage

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(haven)
UNHCR_SSD_2023_FDS_data_main <- read_dta("../00_data/01_input/UNHCR_SSD_2023_FDS_data_v2/UNHCR_SSD_2023_FDS_data_main.dta")
#View(UNHCR_SSD_2023_FDS_data_main)
library(haven)
UNHCR_SSD_2023_FDS_data_member <- read_dta("../00_data/01_input/UNHCR_SSD_2023_FDS_data_v2/UNHCR_SSD_2023_FDS_data_member.dta")

UNHCR_SSD_2023_FDS_data_caregiver <- read_dta("../00_data/01_input/UNHCR_SSD_2023_FDS_data_v2/UNHCR_SSD_2023_FDS_data_caregiver.dta")
#View(UNHCR_SSD_2023_FDS_data_member)

UNHCR_SSD_2023_FDS_data_woman <- read_dta("../00_data/01_input/UNHCR_SSD_2023_FDS_data_v2/UNHCR_SSD_2023_FDS_data_woman.dta")

UNHCR_SSD_2023_FDS_data_roster <- read_dta("../00_data/01_input/UNHCR_SSD_2023_FDS_data_v2/UNHCR_SSD_2023_FDS_data_roster.dta")
data = UNHCR_SSD_2023_FDS_data_main %>% 
  dplyr::left_join(UNHCR_SSD_2023_FDS_data_member, by=c("ID")) %>%
  dplyr::rename_with(~ gsub("\\.x$", "", .x)) %>% 
  dplyr::left_join(UNHCR_SSD_2023_FDS_data_caregiver, by=c("ID")) %>%
  dplyr::rename_with(~ gsub("\\.x$", "", .x)) %>% 
  dplyr::left_join(UNHCR_SSD_2023_FDS_data_woman, by=c("ID")) %>%
  dplyr::rename_with(~ gsub("\\.x$", "", .x)) %>% 
  dplyr::left_join(UNHCR_SSD_2023_FDS_data_roster %>% dplyr::filter(HH_03==1), by=c("ID")) %>%
  dplyr::rename_with(~ gsub("\\.x$", "", .x)) %>%
  mutate(across(where(is.numeric),  ~ na_if(.x, 998))) %>%
  labelled::to_factor()

data$ageYears <- as.numeric(data$ageYears)


data$Intro_07_1 <- droplevels(data$Intro_07_1)
```


# Legal status and displacement history



```{r nationality, echo=FALSE, message=FALSE, warning=FALSE}
var_needed = c("ID_00",
"Legal_16",
"HH_07b",
"HH_00b_year_random",
"DH_13",
"DH_31a",
"DH_31b",
"DH_31c"
)
gtsummary::set_gtsummary_theme(gtsummary::theme_gtsummary_compact())
data %>% dplyr::select(Intro_07_1, dplyr::all_of(var_needed)) %>% 
gtsummary::tbl_summary(by = Intro_07_1,
                            statistic = list(all_categorical() ~ "{p}",
                                             all_continuous() ~ "{mean}({min},{max})"),
                            missing = "no"   ) %>%   gtsummary::add_p() %>% 
  
  modify_header(label ~ "Covariates")
```


# Demographics

```{r age_pyramid, echo=FALSE, fig.height=6, fig.show='asis', fig.width=12, message=FALSE, warning=FALSE, cache=FALSE, results='asis'}
# Function to create age groups
autocut_all <- function(x) {
  cut(
    x,
    breaks = c(0, 5, 10, 15, 20, 25, 30, 35, 40, 45, 50, 55, 60, 65, 70, 75, 80, 85, Inf),
    right = FALSE,
    include.highest = TRUE
  )
}
# Ensure dataset is a data.table
setDT(UNHCR_SSD_2023_FDS_data_roster)

# Ensure age is numeric
UNHCR_SSD_2023_FDS_data_roster[, ageYears := as.numeric(as.character(ageYears))]

# Create age group

UNHCR_SSD_2023_FDS_data_roster <- UNHCR_SSD_2023_FDS_data_roster[, age_group := autocut_all(ageYears)]

UNHCR_SSD_2023_FDS_data_roster[, age_group := as.factor(age_group )]
UNHCR_SSD_2023_FDS_data_roster[, HH_02 := as.factor(HH_02)]
# Plot pyramid
apyramid::age_pyramid(
  data = UNHCR_SSD_2023_FDS_data_roster[!is.na(age_group) & HH_02 != ""],
  age_group = age_group,
  split_by = HH_02
) +
  ylab("# cases") +
  xlab("Age (years)") +
  scale_fill_manual(
    values = c("#e0fbfc", "#1f4e8c"),
    labels = c("Male", "Female")
  ) +
  labs(fill = "Sex") +
  theme_classic()
```

```{r individual_characteristics, echo=FALSE, fig.height=6, fig.show='asis', fig.width=12, message=FALSE, warning=FALSE, cache=FALSE, results='asis'}
apyramid::age_pyramid(
  data = UNHCR_SSD_2023_FDS_data_roster[!is.na(age_group) & HH_02 != "" & !is.na(Intro_07_1)],
  age_group = age_group,
  split_by = HH_02,
  stack_by = Intro_07_1
) +
  ylab("# cases") +
  xlab("Age (years)") +
  scale_fill_manual(
    values = c("#1f4e8c", "#e0fbfc"),
    labels = c("Host community North","Refugees")
  ) +
  labs(fill = "Population group") +
  theme_classic()
```


```{r householdHead_characteristics, echo=FALSE, message=FALSE, warning=FALSE}
var_needed = c("ageYears",
               "HH_02",
               "HH_08",
               "HH_09",
               "HH_21",
               "HH_25"
               #"Intro_08"
)
gtsummary::set_gtsummary_theme(gtsummary::theme_gtsummary_compact())
data$HH_09 = as.numeric(data$HH_09)
Hmisc::label(data$HH_09) <- "Age of first married/formed a non-formal union"
data %>%  dplyr::select(Intro_07_1, dplyr::all_of(var_needed)) %>%
  gtsummary::tbl_summary(by = Intro_07_1,
                         statistic = list(all_categorical() ~ "{p}",
                                          all_continuous() ~ "{mean}({min},{max})"),
                         missing = "no"   ) %>%   gtsummary::add_p() %>% 
  
  modify_header(label ~ "Household head characteristics")
```






# Social environment


```{r mobility, echo=FALSE, message=FALSE, warning=FALSE}
var_needed = c("MovFree01b",
"HH20",
"AS01_1",
"AS01_2",
"AS01_3a",
"AS01_3b",
"AS01_4",
"AS01_5"
)
gtsummary::set_gtsummary_theme(gtsummary::theme_gtsummary_compact())
data %>% dplyr::select(Intro_07_1, dplyr::all_of(var_needed)) %>% 
gtsummary::tbl_summary(by = Intro_07_1,
                            statistic = list(all_categorical() ~ "{p}",
                                             all_continuous() ~ "{mean}({min},{max})"),
                            missing = "no"   ) %>%   gtsummary::add_p() %>% 
  
  modify_header(label ~ "Covariates")
```



# Basic needs

## Water, sanitation, and hygiene (WASH)


```{r drinking_water, echo=FALSE, message=FALSE, warning=FALSE}
var_needed = c("BD11",
"HH12",
"HW6"
)

gtsummary::set_gtsummary_theme(gtsummary::theme_gtsummary_compact())
data %>% dplyr::select(Intro_07_1, dplyr::all_of(var_needed)) %>% 
gtsummary::tbl_summary(by = Intro_07_1,
                            statistic = list(all_categorical() ~ "{p}",
                                             all_continuous() ~ "{mean}({min},{max})"),
                            missing = "no"   ) %>%   gtsummary::add_p() %>% 
  
  modify_header(label ~ "Covariates")
```



## Clean energy


```{r electricity, echo=FALSE, message=FALSE, warning=FALSE}
var_needed = c("HC04",
"HC02"
)
gtsummary::set_gtsummary_theme(gtsummary::theme_gtsummary_compact())
data %>% dplyr::select(Intro_07_1, dplyr::all_of(var_needed)) %>% 
gtsummary::tbl_summary(by = Intro_07_1,
                            statistic = list(all_categorical() ~ "{p}",
                                             all_continuous() ~ "{mean}({min},{max})"),
                            missing = "no"   ) %>%   gtsummary::add_p() %>% 
  
  modify_header(label ~ "Covariates")
```



## Shelter

```{r improved_shelter, echo=FALSE, message=FALSE, warning=FALSE}
var_needed = c("HH01",
"HH02",
"HH06"
)
gtsummary::set_gtsummary_theme(gtsummary::theme_gtsummary_compact())
data %>% dplyr::select(Intro_07_1, dplyr::all_of(var_needed)) %>% 
gtsummary::tbl_summary(by = Intro_07_1,
                            statistic = list(all_categorical() ~ "{p}",
                                             all_continuous() ~ "{mean}({min},{max})"),
                            missing = "no"   ) %>%   gtsummary::add_p() %>% 
  
  modify_header(label ~ "Covariates")
```



## Health

```{r access_to_health_care, echo=FALSE, message=FALSE, warning=FALSE}
var_needed = c("H29",
"MN01",
"MN02",
"Dis_01",
"Dis_04",
"Dis_07",
"Dis_10",
"Dis_13",
"Dis_16"
)
gtsummary::set_gtsummary_theme(gtsummary::theme_gtsummary_compact())
data %>% dplyr::select(Intro_07_1, dplyr::all_of(var_needed)) %>% 
gtsummary::tbl_summary(by = Intro_07_1,
                            statistic = list(all_categorical() ~ "{p}",
                                             all_continuous() ~ "{mean}({min},{max})"),
                            missing = "no"   ) %>%   gtsummary::add_p() %>% 
  
  modify_header(label ~ "Covariates")
```

## Maternal and child health


```{r skilled_birth_attendance, echo=FALSE, message=FALSE, warning=FALSE}
var_needed = c(
"BA10A",
"BA10B",
"BA10D",
"BA10E",
"BA10F",
"BA10G",
"BA10X",
"BA6_1",
"BA17",
"MV1",
"MV1a",
"MV2",
"MV3",
"MV4",
"MV6",
"MV7",
"MV8",
"MV9",
"MV10",
"MV11",
"MV12A",
"MV12B",
"MV12C",
"MV12D",
"MV12E",
"MV12F",
"MV12OT",
"MV12DK"
)
gtsummary::set_gtsummary_theme(gtsummary::theme_gtsummary_compact())
data %>% dplyr::select(Intro_07_1, dplyr::all_of(var_needed)) %>% 
gtsummary::tbl_summary(by = Intro_07_1,
                            statistic = list(all_categorical() ~ "{p}",
                                             all_continuous() ~ "{mean}({min},{max})"),
                            missing = "no"   ) %>%   gtsummary::add_p() %>% 
  
  modify_header(label ~ "Covariates")
```




## Education

```{r HHschool_completion, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
var_needed = c("HH_Educ06",
"HH_Educ07",
"HH_Educ10",
"HH_Educ15a",
"HH_Educ15b",
"HH_Educ16",
"HH_Educ17",
"HH_Educ18",
"HH_Educ23"
)
gtsummary::set_gtsummary_theme(gtsummary::theme_gtsummary_compact())
UNHCR_SSD_2023_FDS_data_roster %>% dplyr::filter(HH_03==1) %>% dplyr::select(Intro_07_1, dplyr::all_of(var_needed)) %>% 
gtsummary::tbl_summary(by = Intro_07_1,
                            statistic = list(all_categorical() ~ "{p}",
                                             all_continuous() ~ "{mean}({min},{max})"),
                            missing = "no"   ) %>%   gtsummary::add_p() %>% 
  
  modify_header(label ~ "Househould head education status")
```


```{r school_completion, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
var_needed = c("HH_Educ06",
"HH_Educ07",
"HH_Educ10",
"HH_Educ15a",
"HH_Educ15b",
"HH_Educ16",
"HH_Educ17",
"HH_Educ18",
"HH_Educ23"
)
gtsummary::set_gtsummary_theme(gtsummary::theme_gtsummary_compact())
UNHCR_SSD_2023_FDS_data_roster %>% dplyr::select(Intro_07_1, dplyr::all_of(var_needed)) %>% 
gtsummary::tbl_summary(by = Intro_07_1,
                            statistic = list(all_categorical() ~ "{p}",
                                             all_continuous() ~ "{mean}({min},{max})"),
                            missing = "no"   ) %>%   gtsummary::add_p() %>% 
  
  modify_header(label ~ "Househould members education status")
```



## Food and Nutrition

```{r echo=FALSE, message=FALSE, warning=FALSE}
var_needed = c("Food00",
"Food03a",
"Food09a",
"Food10a",
"Food11a"
)
gtsummary::set_gtsummary_theme(gtsummary::theme_gtsummary_compact())
data %>% dplyr::select(Intro_07_1, dplyr::all_of(var_needed)) %>% 
gtsummary::tbl_summary(by = Intro_07_1,
                            statistic = list(all_categorical() ~ "{p}",
                                             all_continuous() ~ "{mean}({min},{max})"),
                            missing = "no"   ) %>%   gtsummary::add_p() %>% 
  
  modify_header(label ~ "Covariates")
```


## Social protection

```{r echo=FALSE, message=FALSE, warning=FALSE}
var_needed = c("ScProtec01"
)
gtsummary::set_gtsummary_theme(gtsummary::theme_gtsummary_compact())
data %>% dplyr::select(Intro_07_1, dplyr::all_of(var_needed)) %>% 
gtsummary::tbl_summary(by = Intro_07_1,
                            statistic = list(all_categorical() ~ "{p}",
                                             all_continuous() ~ "{mean}({min},{max})"),
                            missing = "no"   ) %>%   gtsummary::add_p() %>% 
  
  modify_header(label ~ "Covariates")
```


# Livelihoods

## Employment

```{r echo=FALSE, message=FALSE, warning=FALSE}

var_needed = c("EMP01"
)
gtsummary::set_gtsummary_theme(gtsummary::theme_gtsummary_compact())
data %>% dplyr::select(Intro_07_1, dplyr::all_of(var_needed)) %>% 
gtsummary::tbl_summary(by = Intro_07_1,
                            statistic = list(all_categorical() ~ "{p}",
                                             all_continuous() ~ "{mean}({min},{max})"),
                            missing = "no"   ) %>%   gtsummary::add_p() %>% 
  
  modify_header(label ~ "Covariates")
```


## Child Labour

```{r echo=FALSE, message=FALSE, warning=FALSE}
var_needed = c("child_labour",
  "CL01_1",
"CL01_2",
"CL01_3",
"CL01_4",
"CL02"
)
data = data %>% 
  dplyr::mutate(child_labour = ifelse(CL01_1 == "yes"| CL01_2 == "yes" | CL01_3 == "yes"| CL01_4 == "yes",1,0))
gtsummary::set_gtsummary_theme(gtsummary::theme_gtsummary_compact())
data %>% dplyr::select(Intro_07_1, dplyr::all_of(var_needed)) %>% 
gtsummary::tbl_summary(by = Intro_07_1,
                            statistic = list(all_categorical() ~ "{p}",
                                             all_continuous() ~ "{mean}({min},{max})"),
                            missing = "no"   ) %>%   gtsummary::add_p() %>% 
  
  modify_header(label ~ "Covariates")
```


## Financial services


```{r echo=FALSE, message=FALSE, warning=FALSE}
var_needed = c("FI01",
"House20D"
)
gtsummary::set_gtsummary_theme(gtsummary::theme_gtsummary_compact())
data %>% dplyr::select(Intro_07_1, dplyr::all_of(var_needed)) %>% 
gtsummary::tbl_summary(by = Intro_07_1,
                            statistic = list(all_categorical() ~ "{p}",
                                             all_continuous() ~ "{mean}({min},{max})"),
                            missing = "no"   ) %>%   gtsummary::add_p() %>% 
  
  modify_header(label ~ "Covariates")
```




## Land and property

```{r echo=FALSE, message=FALSE, warning=FALSE}
var_needed = c("Land01",
"Land02"
)
gtsummary::set_gtsummary_theme(gtsummary::theme_gtsummary_compact())
data %>% dplyr::select(Intro_07_1, dplyr::all_of(var_needed)) %>% 
gtsummary::tbl_summary(by = Intro_07_1,
                            statistic = list(all_categorical() ~ "{p}",
                                             all_continuous() ~ "{mean}({min},{max})"),
                            missing = "no"   ) %>%   gtsummary::add_p() %>% 
  
  modify_header(label ~ "Covariates")
```



## Shocks

```{r shocks, echo=FALSE, message=FALSE, warning=FALSE}
var_needed = c("ExpShock01"
)
gtsummary::set_gtsummary_theme(gtsummary::theme_gtsummary_compact())
data %>% dplyr::select(Intro_07_1, dplyr::all_of(var_needed)) %>% 
gtsummary::tbl_summary(by = Intro_07_1,
                            statistic = list(all_categorical() ~ "{p}",
                                             all_continuous() ~ "{mean}({min},{max})"),
                            missing = "no"   ) %>%   gtsummary::add_p() %>% 
  
  modify_header(label ~ "Covariates")
```

## Changes to household income

```{r changes, echo=FALSE, message=FALSE, warning=FALSE}
var_needed = c("CH01a"
)
gtsummary::set_gtsummary_theme(gtsummary::theme_gtsummary_compact())
data %>% dplyr::select(Intro_07_1, dplyr::all_of(var_needed)) %>% 
gtsummary::tbl_summary(by = Intro_07_1,
                            statistic = list(all_categorical() ~ "{p}",
                                             all_continuous() ~ "{mean}({min},{max})"),
                            missing = "no"   ) %>%   gtsummary::add_p() %>% 
  
  modify_header(label ~ "Covariates")
```


## Subjective well-being

```{r swellbeing, echo=FALSE, message=FALSE, warning=FALSE}
var_needed = c("W02_1",
"W01_1",
"W01_2"
)
gtsummary::set_gtsummary_theme(gtsummary::theme_gtsummary_compact())
data %>% dplyr::select(Intro_07_1, dplyr::all_of(var_needed)) %>% 
gtsummary::tbl_summary(by = Intro_07_1,
                            statistic = list(all_categorical() ~ "{p}",
                                             all_continuous() ~ "{mean}({min},{max})"),
                            missing = "no"   ) %>%   gtsummary::add_p() %>% 
  
  modify_header(label ~ "Covariates")
```

# Assistance programme

```{r assistance, echo=FALSE, message=FALSE, warning=FALSE}
var_needed = c("ScProtec1_ScProtec03",
"ScProtec1_ScProtec04",
"ScProtec1_ScProtec05",
"ScProtec1_ScProtec06",
"ScProtec1_ScProtec07"
)
   
gtsummary::set_gtsummary_theme(gtsummary::theme_gtsummary_compact())
data %>% dplyr::select(Intro_07_1, dplyr::all_of(var_needed)) %>% 
gtsummary::tbl_summary(by = Intro_07_1,
                            statistic = list(all_categorical() ~ "{p}",
                                             all_continuous() ~ "{mean}({min},{max})"),
                            missing = "no"   ) %>%   gtsummary::add_p() %>% 
  
  modify_header(label ~ "Covariates")
```

