---
title: "The multi-dimensional deprivation index (MDDI)"
author: "Data4Decision team - data source : Forced Displacement Survey 2023- South Sudan"
date: "`r format(Sys.time(), '%d %B %Y')`"
always_allow_html: yes
header-includes:
  - \usepackage{booktabs}
  - \usepackage{float}
  - \usepackage{color}
  - \usepackage{longtable}
  - \usepackage{multirow}
output: 
  # prettydoc::html_pretty:
  #    theme: cayman
  officedown::rdocx_document:
    tables:
      style: Table
      layout: autofit
      width: 1
      topcaption: true
      tab.lp: 'tab:'
      caption:
        style: Table Caption
        pre: 'Tab '
        sep: ': '
        tnd: 0
        tns: '-'
        fp_text: !expr officer::fp_text_lite(bold = TRUE)
      conditional:
        first_row: true
        first_column: false
        last_row: false
        last_column: false
        no_hband: false
        no_vband: true
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: true
    toc_depth: 3
    number_sections: true
    theme: lumen
  pdf_document:
    keep_tex: no
  word_document:
    fig_width: 6.5
    fig_height: 3.5
    fig_caption: true
#html_notebook: default
---
  
\newpage

```{r echo=FALSE, message=FALSE, warning=FALSE}

UNHCR_SSD_2023_FDS_data_main <- read_dta("../00_data/01_input/UNHCR_SSD_2023_FDS_data_v2/UNHCR_SSD_2023_FDS_data_main.dta") %>%
   labelled::to_factor()
#View(UNHCR_SSD_2023_FDS_data_main)

UNHCR_SSD_2023_FDS_data_roster <- read_dta("../00_data/01_input/UNHCR_SSD_2023_FDS_data_v2/UNHCR_SSD_2023_FDS_data_roster.dta") %>%
   labelled::to_factor()
data = UNHCR_SSD_2023_FDS_data_main



data$Intro_07_1 <- droplevels(data$Intro_07_1)
```


## Food Consumption Score (FCS)

```{r echo=FALSE, message=FALSE, warning=FALSE}
data = data %>% 
  dplyr::mutate(
    FCSStap=Food_div1,
    FCSPulse=Food_div2,
    FCSDairy=Food_div3,
    FCSPr=Food_div4,
    FCSVeg=Food_div5,
    FCSFruit=Food_div6,
    FCSFruitOrg=Food_div61,
    FCSFat=Food_div7,
    FCSSugar=Food_div8,
    FCSCond=Food_div9
  )



# Label relevant FCS variables ------------------------------------------------# 

var_label(data$FCSStap)   <- "Consumption over the past 7 days: cereals, grains and tubers"
var_label(data$FCSPulse)  <- "Consumption over the past 7 days: pulses"
var_label(data$FCSDairy)  <- "Consumption over the past 7 days: dairy products"
var_label(data$FCSPr)     <- "Consumption over the past 7 days: meat, fish and eggs"
var_label(data$FCSVeg)    <- "Consumption over the past 7 days: vegetables"
var_label(data$FCSFruit)  <- "Consumption over the past 7 days: fruit"
var_label(data$FCSFat)    <- "Consumption over the past 7 days: fat and oil"
var_label(data$FCSSugar)  <- "Consumption over the past 7 days: sugar or sweets:"
var_label(data$FCSCond)   <- "Consumption over the past 7 days: condiments or spices"

# Calculate FCS ---------------------------------------------------------------# 
data <- data %>% mutate(FCS = (FCSStap  * 2)   + 
                              (FCSPulse * 3)   +
                              (FCSPr    * 4)   +
                              (FCSDairy * 4)   + 
                               FCSVeg          +
                               FCSFruit        +
                              (FCSFat   * 0.5) +
                              (FCSSugar * 0.5))
var_label(data$FCS) <- "Food Consumption Score"
data <- data %>% mutate(FCSCat21 = case_when(
                             FCS <= 21 ~ 1,
                             between(FCS, 21.5, 35) ~ 2, 
                             FCS > 35 ~ 3),
                        FCSCat28 = case_when(
                             FCS <= 28 ~ 1, 
                             between(FCS, 28.5, 42) ~ 2, 
                             FCS > 42 ~ 3))

val_lab(data$FCSCat21) = num_lab("
             1 Poor
             2 Borderline
             3 Acceptable
                                 ")
var_label(data$FCSCat21) <- "FCS Categories: 21/35 thresholds"

val_lab(data$FCSCat28) = num_lab("
             1 Poor
             2 Borderline
             3 Acceptable
                                 ")
var_label(data$FCSCat28) <- "FCS Categories: 28/42 thresholds"
```


```{r echo=FALSE, message=FALSE, warning=FALSE}
var_needed = c("FCS",
               "FCSCat21",
               "FCSCat28"
)
gtsummary::set_gtsummary_theme(gtsummary::theme_gtsummary_compact())
data %>% dplyr::select(Intro_07_1, dplyr::all_of(var_needed)) %>% 
gtsummary::tbl_summary(by = Intro_07_1,
                            statistic = list(all_categorical() ~ "{p}",
                                             all_continuous() ~ "{mean}({min},{max})"),
                            missing = "no"   ) %>%   gtsummary::add_p() %>% 
  
  modify_header(label ~ "Covariates")
```

# Livelihood coping stategies index


```{r echo=FALSE, message=FALSE, warning=FALSE}
data = data %>% 
  dplyr::mutate(
    LhCSIStress1= ifelse(Food10_1p=="to buy food",1,0),
    LhCSIStress2= ifelse(Food10_1h=="to buy food",1,0),
    LhCSIStress3= ifelse(Food10_1z=="to buy food",1,0),
    LhCSIStress4= ifelse(Food10_1k=="to buy food",1,0),
    LhCSICrisis1= ifelse(Food10_1g=="to buy food",1,0),
    LhCSICrisis2= ifelse(Food10_1i=="to buy food",1,0),
    LhCSICrisis3= ifelse(Food10_1m=="to buy food",1,ifelse(Food10_1l=="to buy food",1,0)),
    LhCSIEmergency1= ifelse(Food10_1q=="to buy food",1,0),
    LhCSIEmergency2= ifelse(Food10_1b=="to buy food",1,0),
    LhCSIEmergency3= ifelse(Food10_1n=="to buy food",1,ifelse(Food10_1c=="to buy food",1,0))
    

)

#stress
data <- data %>% mutate(stress_coping_FS = case_when(
  LhCSIStress1 == 1 ~ 1,
  LhCSIStress2 == 1 ~ 1,
  LhCSIStress3 == 1 ~ 1,
  LhCSIStress4 == 1 ~1,
  TRUE ~ 0))
var_label(data$stress_coping_FS) <- "Did the HH engage in stress coping strategies"
#Crisis
data <- data %>% mutate(crisis_coping_FS = case_when(
  LhCSICrisis1 == 1 ~ 1,
  LhCSICrisis2 == 1 ~ 1,
  LhCSICrisis3 == 1 ~ 1,
  TRUE ~ 0))
var_label(data$crisis_coping_FS) <- "Did the HH engage in crisis coping strategies"
#Emergency
data <- data %>% mutate(emergency_coping_FS = case_when(
  LhCSIEmergency1 == 1 ~ 1,
  LhCSIEmergency2 == 1 ~ 1,
  LhCSIEmergency3 == 1 ~ 1,
  TRUE ~ 0))
var_label(data$emergency_coping_FS) <- "Did the HH engage in emergency coping strategies"

#calculate Max_coping_behaviour
data <- data %>% mutate(Max_coping_behaviourFS = case_when(
  is.na(emergency_coping_FS) & is.na(crisis_coping_FS) & is.na(stress_coping_FS) ~ NA_real_,
  emergency_coping_FS == 1 ~ 4,
  crisis_coping_FS == 1 ~ 3,
  stress_coping_FS == 1 ~ 2,
  TRUE ~ 1))
var_label(data$Max_coping_behaviourFS) <- "Summary of asset depletion"
val_lab(data$Max_coping_behaviourFS) = num_lab("
             1 HH not adopting coping strategies
             2 Stress coping strategies
             3 Crisis coping strategies
             4 Emergencies coping strategies
")
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
var_needed = c("Max_coping_behaviourFS"
)
gtsummary::set_gtsummary_theme(gtsummary::theme_gtsummary_compact())
data %>% dplyr::select(Intro_07_1, dplyr::all_of(var_needed)) %>% 
gtsummary::tbl_summary(by = Intro_07_1,
                            statistic = list(all_categorical() ~ "{p}",
                                             all_continuous() ~ "{mean}({min},{max})"),
                            missing = "no"   ) %>%   gtsummary::add_p() %>% 
  
  modify_header(label ~ "Covariates")
```



## Reduced Consumption Strategies Index (rCSI)

```{r echo=FALSE, message=FALSE, warning=FALSE}
data <- data %>% mutate(
  rCSILessQlty= ifelse(Food02a=="yes",4,0),
  rCSIBorrow= ifelse(Food05a=="yes",4,0),
  rCSIMealNb= ifelse(Food08a=="yes",4,0),
  rCSIMealSize= ifelse(Food06a=="yes",4,0),
  rCSIMealAdult= ifelse(Food07a=="yes",4,0)

)


# Label rCSI relevant variables -----------------------------------------------#

var_label(data$rCSILessQlty)  <- "Relied on less preferred and less expensive food"
var_label(data$rCSIBorrow)    <- "Borrowed food or relied on help from a relative or friend"
var_label(data$rCSIMealNb)    <- "Reduce number of meals eaten in a day"
var_label(data$rCSIMealSize)  <- "Limit portion size of meals at meal times"
var_label(data$rCSIMealAdult) <- "Restricted consumption by adults for small children to eat"

# Calculate rCSI --------------------------------------------------------------# 

data <- data %>% mutate(rCSI = rCSILessQlty + 
                              rCSIBorrow * 2 + 
                              rCSIMealNb + 
                              rCSIMealSize + 
                              rCSIMealAdult * 3)

data$rCSI <- as.numeric(data$rCSI)
var_label(data$rCSI)          <- "Reduced coping strategies index (rCSI)"
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
var_needed = c("rCSI")
gtsummary::set_gtsummary_theme(gtsummary::theme_gtsummary_compact())
data %>% dplyr::select(Intro_07_1, dplyr::all_of(var_needed)) %>% 
gtsummary::tbl_summary(by = Intro_07_1,
                       type = list(rCSI ~ "continuous"),
                            statistic = list(all_categorical() ~ "{p}",
                                             all_continuous() ~ "{mean}({min},{max})"),
                            missing = "no"   ) %>%   gtsummary::add_p() %>% 
  
  modify_header(label ~ "Covariates")
```


# Multi-dimensional deprivation index (MDDI) is a composite score of poverty or deprivation

## FOOD DIMENSION

### Food Consumption Score

```{r echo=FALSE, message=FALSE, warning=FALSE}
#HH with unacceptable food consumption
data = data %>% 
  dplyr::mutate(MDDI_food1 = ifelse(!is.na(FCSCat28) & (FCSCat28==1 | FCSCat28==2),1,0))

var_label(data$MDDI_food1)   <- "HH with unacceptable food consumption"
```



### Reduced Consumption Strategies Index

```{r echo=FALSE, message=FALSE, warning=FALSE}
#HH with high level of consumption coping strategies
data = data %>% 
  dplyr::mutate(MDDI_food2 = ifelse(!is.na(rCSI) & rCSI>18,1,0))

var_label(data$MDDI_food2)   <- "HH with high level of consumption coping strategies"
```



## EDUCATION DIMENSION

### At least one school age children (6-17) (adjust to country context) not attending school in the last 6 months

```{r echo=FALSE, message=FALSE, warning=FALSE}
#HH with at least one school-age children not attending school
UNHCR_SSD_2023_FDS_data_roster$ageYears <- as.numeric(UNHCR_SSD_2023_FDS_data_roster$ageYears)
UNHCR_SSD_2023_FDS_data_roster = UNHCR_SSD_2023_FDS_data_roster %>%
  dplyr::mutate(HHNoSchool = ifelse((ageYears>=6 & ageYears<=17) & !is.na(HH_Educ06),1,0))

df = UNHCR_SSD_2023_FDS_data_roster %>% 
  dplyr::group_by(ID) %>% 
  dplyr::mutate(HHNoSchoolNum = sum(HHNoSchool,na.rm = T)) %>% 
  dplyr::ungroup() %>% 
  dplyr::mutate(MDDI_edu1 = ifelse(HHNoSchoolNum>0,1,0)) %>% 
  dplyr::select(ID,MDDI_edu1) %>% 
  dplyr::distinct()

data = data %>% 
  dplyr::left_join(df,by="ID")

var_label(data$MDDI_edu1)   <- "HH with at least one school-age children not attending school"
```


## HEALTH DIMENSION

### Medical treatment - Did household members being chronically or acutely ill receive medical attention while sick?

```{r echo=FALSE, message=FALSE, warning=FALSE}
#HH with at least one member did not receive medical treatment while sick

```

### Number of sick or disabled people > 1 or >50% of household members

```{r echo=FALSE, message=FALSE, warning=FALSE}
#HH with more than half members or more than one member sick or disabled
data = data %>% 
  dplyr::mutate(
    dis_sum = rowSums(across(c(Dis_01, Dis_04, Dis_07, Dis_10, Dis_13, Dis_16), ~ .x == 1)),
 MDDI_health2 = ifelse(dis_sum>=1,1,0))

var_label(data$MDDI_health2)   <- "HH with more than half members or more than one member sick or disabled"
```


## SHELTER DIMENSION

### Source of energy for cooking - HH uses solid fuels for cooking

```{r echo=FALSE, message=FALSE, warning=FALSE}
#HH with no improved energy source for cooking
data = data %>% 
  dplyr::mutate(MDDI_shelter1 = ifelse(!is.na(HC02) & (HC02 == "traditional solid fuel stove (non-manufactured)"| HC02 == "three stone stove / open fire" | HC02 == "Build stove with bricks and/or mud"),1,0))

var_label(data$MDDI_shelter1)   <- "HH with no improved energy source for cooking"
```

### Source of energy for lighting - HH has no electricity

```{r echo=FALSE, message=FALSE, warning=FALSE}
#HH with not improved source of energy for lighting
data = data %>% 
  dplyr::mutate(MDDI_shelter2 = ifelse(!is.na(HL01) & HL01 == "no",1,0))

var_label(data$MDDI_shelter2)   <- "HH with not improved source of energy for lighting"
```


### Crowding Index - (Number of HH members/Number of rooms (excluding kitchen, corridors))>3

```{r echo=FALSE, message=FALSE, warning=FALSE}
#HH with at least 3 HH members sharing one room to sleep

df = UNHCR_SSD_2023_FDS_data_roster %>% 
  dplyr::group_by(ID) %>% 
  dplyr::mutate(HHSize = n()) %>% 
  dplyr::ungroup() %>% 
  dplyr::select(ID,HHSize) %>% 
  dplyr::distinct()
data = data %>% 
  dplyr::left_join(df,by="ID")
data = data %>% 
  dplyr::mutate(crowding=HHSize/HH14,
                MDDI_shelter3 = ifelse(crowding>3,1,0))

var_label(data$MDDI_shelter3)   <- "HH with at least 3 HH members sharing one room to sleep"
```




## WASH DIMENSION

### Toilet Type (not-improved facility)

```{r echo=FALSE, message=FALSE, warning=FALSE}
#HH with not improved toilet facility
data = data %>% 
  dplyr::mutate(MDDI_wash1 = ifelse(!is.na(SAN01) & SAN01 == "no facility / bush / field",1,0))

var_label(data$MDDI_wash1)   <- "HH with not improved toilet facility"
```


### Water source (not-improved source)

```{r echo=FALSE, message=FALSE, warning=FALSE}
#HH with not improved drinking water source
data = data %>% 
  dplyr::mutate(MDDI_wash2 = ifelse(!is.na(BD01) & (BD01 == "unprotected well"| BD01 == "unprotected spring" | BD01 == "surface water (river, dam, lake, pond, stream, canal, irrigation channel)"),1,0))

var_label(data$MDDI_wash2)   <- "HH with not improved drinking water source"
```



## SAFETY DIMENSION

### Safety: HH felt unsafe or suffered violence

```{r echo=FALSE, message=FALSE, warning=FALSE}
#HH with one or more members who felt unsafe or suffered violence
data = data %>% 
  dplyr::mutate(MDDI_safety1 = ifelse(!is.na(HH20) & HH20 == "never feel safe",1,0))

var_label(data$MDDI_safety1)   <- "HH with one or more members who felt unsafe or suffered violence"
```


### Displaced by forced in the last 12 months

```{r echo=FALSE, message=FALSE, warning=FALSE}
#HH displaced by force in the last 12 months
```


```{r echo=FALSE, message=FALSE, warning=FALSE}
var_needed = data %>% 
  dplyr::select(gtsummary::starts_with("MDDI")) %>% 
  names()
gtsummary::set_gtsummary_theme(gtsummary::theme_gtsummary_compact())
data %>% dplyr::select(Intro_07_1, dplyr::all_of(var_needed)) %>% 
gtsummary::tbl_summary(by = Intro_07_1,
                            statistic = list(all_categorical() ~ "{p}",
                                             all_continuous() ~ "{mean}({min},{max})"),
                            missing = "no"   ) %>%   gtsummary::add_p() %>% 
  
  modify_header(label ~ "Covariates")
```



## Calculate deprivation score of each dimension

For each indicator, a positive answer was coded as 1 and a negative answer as 0. Each answer was then multiplied by the indicator weight. Summed together, each household gets a score out of 1, where any value greater than or equal to 0.3 equates to ‘multidimensional poverty’, i.e. deprivation across multiple sectors (‘MD Poor’).

A household is considered deprived in any one dimension for which at least 50% of the indicators are coded as positive answers.

```{r echo=FALSE, message=FALSE, warning=FALSE}
data <- data %>%
  mutate(
    MDDI_food    = MDDI_food1 * 0.5 + MDDI_food2 * 0.5,
    MDDI_edu     = MDDI_edu1 * 1,
    MDDI_health  =  MDDI_health2 * 1,
    MDDI_shelter = MDDI_shelter1 * 1/3 + 
                   MDDI_shelter2 * 1/3 + 
                   MDDI_shelter3 * 1/3,
    MDDI_wash    = MDDI_wash1 * 0.5 + 
                   MDDI_wash2 * 0.5,
    MDDI_safety  = MDDI_safety1 * 1
  )

data$MDDI_food <- as.numeric(data$MDDI_food)
data$MDDI_edu <- as.numeric(data$MDDI_edu)
data$MDDI_health <- as.numeric(data$MDDI_health)
data$MDDI_shelter <- as.numeric(data$MDDI_shelter)
data$MDDI_wash <- as.numeric(data$MDDI_wash)
data$MDDI_safety <- as.numeric(data$MDDI_safety)

var_label(data$MDDI_food)    <- "Deprivation score for food dimension"
var_label(data$MDDI_edu)     <- "Deprivation score for education dimension"
var_label(data$MDDI_health)  <- "Deprivation score for health dimension"
var_label(data$MDDI_shelter) <- "Deprivation score for shelter dimension"
var_label(data$MDDI_wash)    <- "Deprivation score for WASH dimension"
var_label(data$MDDI_safety)  <- "Deprivation score for safety and displacement dimension"



data <- data %>%
  mutate(
    MDDI = (MDDI_food + MDDI_edu + MDDI_health + MDDI_shelter + MDDI_wash + MDDI_safety) / 6
  )

var_label(data$MDDI) <- "MDDI score"



```

```{r echo=FALSE, message=FALSE, warning=FALSE}

var_needed = c("MDDI_food", "MDDI_edu", "MDDI_health", "MDDI_shelter", "MDDI_wash", "MDDI_safety","MDDI")
gtsummary::set_gtsummary_theme(gtsummary::theme_gtsummary_compact())
data %>% dplyr::select(Intro_07_1, dplyr::all_of(var_needed)) %>% 
gtsummary::tbl_summary(by = Intro_07_1,
                            statistic = list(all_categorical() ~ "{p}",
                                             all_continuous() ~ "{mean}({min},{max})"),
                            missing = "no"   ) %>%   gtsummary::add_p() %>% 
  
  modify_header(label ~ "Covariates")
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
#Calculate MDDI Incidence
data <- data %>%
  mutate(
    MDDI_poor_severe = ifelse(MDDI >= 0.50,1,0)
  )

var_label(data$MDDI_poor_severe) <- "MDDI Incidence – severe deprivation"

data <- data %>%
  mutate(
    MDDI_poor = ifelse(MDDI >= 0.3,1,0)
  )

var_label(data$MDDI_poor) <- "MDDI Incidence"

# Apply labels to variables
data <- data %>%
  mutate(
    MDDI_poor_severe = factor(MDDI_poor_severe, levels = c(0, 1), labels = c("HH is not deprived", "HH is deprived")),
    MDDI_poor = factor(MDDI_poor, levels = c(0, 1), labels = c("HH is not deprived", "HH is deprived"))
  )
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
#Calculate the Average MDDI Intensity
data <- data %>%
  mutate(
    MDDI_intensity = ifelse(MDDI_poor == "HH is deprived", MDDI, 0)
  )

var_label(data$MDDI_intensity) <- "Average MDDI Intensity"

#Calculate Combined MDDI
data <- data %>%
  mutate(
    MDDI_combined = case_when(
      MDDI_poor == "HH is deprived" ~ 1 * MDDI_intensity,
      MDDI_poor == "HH is not deprived" ~ 0,
      TRUE ~ NA_real_  # Preserve NA for missing MDDI_poor
    )
  )

var_label(data$MDDI_combined) <- "Combined MDDI"

```

# Calculate MDDI-related measures

```{r echo=FALSE, message=FALSE, warning=FALSE}
var_needed = c("MDDI_poor", "MDDI_poor_severe", "MDDI")
var_label(data$MDDI_poor) <- "MDDI Incidence (thresholds =  0.33)"
var_label(data$MDDI_poor_severe) <- "MDDI Incidence – severe deprivation (thresholds =  0.50)"
gtsummary::set_gtsummary_theme(gtsummary::theme_gtsummary_compact())
data %>% dplyr::select(Intro_07_1, dplyr::all_of(var_needed)) %>% 
gtsummary::tbl_summary(by = Intro_07_1,
                            statistic = list(all_categorical() ~ "{p}",
                                             all_continuous() ~ "{mean}({min},{max})"),
                            missing = "no"   ) %>%   gtsummary::add_p() %>% 
  
  modify_header(label ~ "MDDI")
```



```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
data = data %>% 
  dplyr::mutate(mddi_edu_1 = ifelse(HH_Educ07==2 & (ageYears<18 | ageYears>=6),1,0), #Not all school age children (6-17) attending school
                mddi_edu_2 = ifelse(HH_Educ16==3 & (ageYears<18 | ageYears>=6),1,0), #education interruption:High costs
             mddi_health_1= ifelse(MN01==2,1,0),
             mddi_health_2 = ifelse(H29==1,1,0),
             mddi_food_1 = ifelse(Food00==1,1,0),
             mddi_food_2 = ifelse(Food11a==1,1,0)) 

```


